Preliminary data analysis

### 2015-05-04 descriptive_analysis.Rmd
Added ordered logistic model results for the first two questions.

```{r, message=FALSE, eval=FALSE}

library(MASS)

d$cPolicy <- ifelse(d$PolicyAdoptedAtPublication=="After JDAP", .5, -.5)
d$cDate <- as.yearmon(d$PublicationDate) %>% as.double %>% scale(., scale=F)

summary(d$cDate)

m <- polr(Q4 ~ 1 + cPolicy + cDate, data=d, Hess=TRUE)

summary(m)

>>>>

### Inferential statistics

```{r, message=FALSE, eval=TRUE, echo=FALSE}

library(MASS)
library(zoo)
library(dplyr)
library(knitr)

d$cAfterPolicyChange <- ifelse(d$PolicyAdoptedAtPublication=="After JDAP", 1, 0)
d$cNonJDAPJournal <- ifelse(d$PolicyAdoptedAtPublication=="No JDAP", 1, 0)
d$cDate <- as.double(as.yearmon(d$PublicationDate)) %>% scale(., scale=F)

d$respQ4 <- ordered(d$Q4, levels=rev(levels(d$Q4)))
d$respQ5 <- ordered(d$Q5, levels=rev(levels(d$Q5)))

m4 <- polr(respQ4 ~ 1 + cAfterPolicyChange + cNonJDAPJournal + cDate, data=d , Hess=TRUE)
# summary(m4)
m4.justchange <- polr(respQ4 ~ 1 + cAfterPolicyChange, data=d , Hess=TRUE)
m4.nochange <- polr(respQ4 ~ 1 + cNonJDAPJournal + cDate, data=d , Hess=TRUE)
suppressMessages( ci4 <- confint(m4) )
table4 <- exp(cbind(OR = coef(m4), ci4)) %>% cbind(., t=coef(summary(m4))[1:nrow(.),'t value'])
#table4 <- cbind(lOR = coef(m4), ci4) %>% cbind(., t=coef(summary(m4))[1:nrow(.),'t value'])

#anova(m4, m4.justchange, m4.nochange)

m5 <- polr(respQ5 ~ 1  + cAfterPolicyChange + cNonJDAPJournal + cDate, data=d, Hess=TRUE)
m5.justchange <- polr(Q5 ~ 1 + cAfterPolicyChange , data=d, Hess=TRUE)
m5.nochange <- polr(Q5 ~ 1 + cNonJDAPJournal + cDate, data=d, Hess=TRUE)
# summary(m5)
suppressMessages( ci5 <- confint(m5) )
table5 <- exp(cbind(OR = coef(m5), ci5)) %>% cbind(., t=coef(summary(m5))[1:nrow(.),'t value'])

#anova(m5, m5.justchange, m5.nochange)
```

The potential effect of JDAP adoption on authors' attitudes is might be confounded by two other factors. Firstly, groups of researchers publishing in journals which are not planning to adopt the JDAP (or at least have not done so yet) may inherently differ from groups publishing in journals which were open to adopting the JDAP. Secondly, attitudes regarding data archiving may improve over time.

In order to test for an effect of policy change on attitudes while controlling for potential confounding effects, we used an ordinal logistic regression model with three predictors: (i) journal type _(cNonJDAPJournal)_, (ii) (centered) publication date in years _(cDate)_, and (iii) a variable indicating whether the article was published after the the journal adopted a data archiving policy _(cAfterPolicyChange)_.

The following table shows the estimated coefficients and 95\%-confidence intervals (transformed to odds ratios) for responses to *question 1*. Effects are statistically significant at the $\alpha=0.05$ level when the transformed confidence intervals do not contain $1$. Confidence intervals were obtained by profile likelihood. According to the model estimates, the introduction of the JDAP data archiving policy increases the odds of a positive response regarding *data sharing* by factor $1.36$ (CI=$[1.16; 1.60]$), while the remaining two predictors do not show any statistically significant effects.
```{r, message=FALSE, eval=TRUE, echo=FALSE}
options(digits=2)
table4
```

The following table shows the estimated coefficients and 95\%-confidence intervals (transformed to odds ratios) for responses to *question 2*. According to the model estimates, the introduction of the JDAP data archiving policy increases the odds of a positive response regarding *data archiving* by factor $1.57$ (CI=$[1.33; 1.90]$), while the remaining two predictors do not show any statistically significant effects.

```{r, message=FALSE, eval=TRUE, echo=FALSE}
options(digits=2)
table5
```
